Collaborative Visual Analysis on the Web with RCloud

### 2015-06-16 system.tex
yes

The starting point in our design was to take advantage of the current
infrastructure of web software as much as possible. HTTP, despite its
deficiencies, has become the lingua franca of
distributed interprocess communication. This meant that if our system
could speak HTTP natively, then it would be possible to provide a low-friction
path from development of an interactive data analysis script to an
automated \emph{web service}, essentially a remote function call over
the web, which can be invoked by higher-level tools. An example of
this sort of progression is described in Section~\ref{sec:casestudy}.

@@

The main unit of computation in RCloud is a \emph{notebook}.  A
notebook holds a sequence of \emph{cells}, each of which is a snippet
of code or hypertext in Markdown. As mentioned in
Section~\ref{sec:related}, this is not novel; executable
documents like this are a feature of many other systems,
including Mathematica, IPython and Sage.

Notebooks can be executed one cell at a time during an interactive
editing session, providing a similar experience to traditional
read-eval-print loops, or can also be executed all at once, providing
an experience similar to running a shell script.

One of the main contributions of RCloud is the notion that notebooks are
``always deployed''. In other words, the most recent version of a
notebook is immediately available to all other users of the system as
the notebook is being developed. Another way to describe this is that
RCloud lacks a ``save'' button: any notebook cell that runs is always
associated to a notebook version serialized to disk. Although one of
our interview subjects reported that this sometimes leads to an
excessively fine-grained sequence of versions (see
Section~\ref{sec:interviews}), we believe that the alternative of
losing automatic sharing is worse (and see
Section~\ref{sec:interviews} again for users reporting their
satisfaction over the low-friction shareability).

When we combine immediate deployment with the ability to refer to the
latest version of a notebook by name, we get notebooks that are always live,
but sometimes broken. Because stability is important, we also
allow any previous version of a notebook to be tagged and referenced.
Our notebook scheme is similar in many ways to models like
Jankun-Kelly et al.'s p-set calculus
\cite{Jankun-Kelly:2007:MFV} and VisTrails's version tree
\cite{Callahan:2006:VVM}, where every change in the state of the
system is tracked.

@@

One side advantage of centralizing the execution and storage of
notebooks is that it becomes feasible to collect usage information
that would otherwise be lost. In the case of social data visualization
platforms, we would like to exploit usage data to help analysts
find content of interest, whether the content is actual source code
or accompanying data. The standard way to achieve this is through
\emph{user-generated curation} and \emph{automatic recommendations}.

Automatic recommendations have become famous in the user experience
provided by companies such as Amazon and Netflix (``If you liked
that film, then you should like this one too''). To compute such
recommendations, we need users to \emph{curate} the collection, by
providing explicit reviews or some other method of indicating interest.
%%
%% In RCloud, reputation and interest are a relationship between
%% \emph{notebooks} and \emph{users}, rather than a relationship between
%% user pairs. We chose this approach because we expect initial 
%% RCloud deployments to have relatively few users, but some users to
%% create many notebooks. Under that assumption, assigning interest
%% to users would not provide sufficiently ``high-resolution'' data.
%%
We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is indicated by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always available, to encourage its use.
%
RCloud today uses only simple counts of stars to measure overall
notebook interest, mostly because standard
recommender system algorithms 
require reasonably large training sets to pay off.
Nevertheless, the starring mechanism is sufficient
to create personalized recommendations~\cite{Hu:2008:CFF}.

Implicit signaling of interest is supported by keeping click-through
counts~\cite{Joachims:2005:AIC} and execution counts. In addition to
these standard techniques of collecting feedback from web search, we
anticipate applying static and dynamic code analysis to infer
fine-grained information about relationships, for example, which
packages and data sets often appear together, in the style of Fast et
al's Codex system.~\cite{Fast:2014:ECS}.

>>>>

Our design takes advantage of existing web software as much as possible.
HTTP, despite its deficiencies, is the lingua franca of distributed
interprocess communication. By speaking HTTP natively, our system
provides a low-friction path from experimental data analysis scripts
to automated \emph{web services}. 
essentially a remote function call over
the web, which can be invoked by higher-level tools. An example of
this sort of progression is described in Section~\ref{sec:casestudy}.

@@

The main unit of computation in RCloud is a \emph{notebook}.  A
notebook holds a sequence of \emph{cells}, each of which is a snippet
of code or hypertext in Markdown. As mentioned in
Section~\ref{sec:related}, executable documents like
this are a feature of previous systems, including Mathematica,
IPython and Sage.

Notebooks can be executed one cell at a time in an interactive
session, similar to traditional read-eval-print
loops, or can all be executed concurrently,
similar to running a shell script.

One of the main contributions of RCloud is the notion that notebooks are
``always deployed''. In other words, the most recent version of a
notebook is immediately available to all other users of the system.
Another way to describe this is that
RCloud lacks a ``save'' button: any notebook cell that runs is always
associated to a notebook version serialized to disk. Although one of
our interview subjects reported that this sometimes leads to an
excessively fine-grained sequence of versions (see
Section~\ref{sec:interviews}), we believe that the alternative of
losing automatic sharing is worse (and see
Section~\ref{sec:interviews} again for users reporting their
satisfaction over the low-friction shareability).

In this design, we get notebooks that are always live,
but sometimes broken. Because stability is important, we also
allow any previous version of a notebook to be tagged and referenced.
Our notebook scheme is similar in many ways to models like
Jankun-Kelly et al.'s p-set calculus
\cite{Jankun-Kelly:2007:MFV} and VisTrails's version tree
\cite{Callahan:2006:VVM}, where every change in the state of the
system is tracked.

@@

One side benefit of centralizing the execution and storage of
notebooks is that it becomes feasible to collect usage information
that might be lost in a conventional environment.
In the case of social data visualization
platforms, we would like to exploit usage data to help analysts
find content of interest, whether the content is actual source code
or accompanying data. The standard way to achieve this is through
\emph{user-generated curation} and \emph{automatic recommendations}.

Automatic recommendations have become famous in the user experience
provided by companies such as Amazon and Netflix (``If you liked
that film, then you should like this one too''). To make such
recommendations, we need users to \emph{curate} the collection, by
providing explicit reviews or some other method of indicating interest.
%%
%% In RCloud, reputation and interest are a relationship between
%% \emph{notebooks} and \emph{users}, rather than a relationship between
%% user pairs. We chose this approach because we expect initial 
%% RCloud deployments to have relatively few users, but some users to
%% create many notebooks. Under that assumption, assigning interest
%% to users would not provide sufficiently ``high-resolution'' data.
%%
We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is indicated by ``starring,'' or
clicking on a button that marks a notebook as interesting.
This makes explicit indication of interest a nearly trivial operation,
always available and easy to use.
%
RCloud today uses only simple counts of stars to measure overall
notebook interest, mostly because standard
recommender system algorithms 
require reasonably large training sets to pay off.
Nevertheless, the starring mechanism is sufficient
to create personalized recommendations~\cite{Hu:2008:CFF}.

Implicit signaling of interest is supported by keeping click-through
counts~\cite{Joachims:2005:AIC} and execution counts. In addition to
these standard techniques, we hope to apply static and dynamic code
analysis to infer fine-grained information about relationships,
for example, which packages and data sets often appear together,
in the style of Fast et al's Codex system.~\cite{Fast:2014:ECS}.
